---
title: "pmartRseq Data Report"
output: word_document
params:
  data: NA
  classes: NA
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=FALSE, eval=TRUE,
                      warning=FALSE, message=FALSE, comment="")
```

```{r setup, include=FALSE}
library(knitr)
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
```

```{r}
# Check for various data objects

# classes <- lapply(data, class)
# classes <- lapply(classes, function(x) x[[1]])
# classes <- unlist(classes)

# info.data <- list()
# info.data$data <- grep("seqData", classes)
# info.data$richness <- grep("richRes", classes)
# info.data$abundance <- grep("abunRes", classes)
# info.data$alphaDiversity <- grep("alphaRes", classes)
# info.data$evenness <- grep("evenRes", classes)
# info.data$jaccard <- grep("jaccard", classes)
# info.data$effectiveSpecies <- grep("effspRes", classes)
# info.data$statistics <- grep("countSTAT_results", classes)
# info.data$indicator <- grep("indspRes", classes)
# 
# # Must only have one item for each class type
# if(any(unlist(lapply(info.data, function(x) length(x))) > 1)){
#   stop(paste("Multiple objects of class ", class(data[[which(unlist(lapply(info.data, length)) > 1)]]), ". This function can only accept one. In the case of multiple data objects, only use the normalized data object.", sep=""))
# }
classes <- params$classes
data <- params$data
# Have to at least have data
if(length(grep("seqData", classes)) == 0){
  stop("Must include data.")
}

```

# Summary of data

```{r}
norm_data <- data$data
```

This `r attr(norm_data, "data_info")$data_type` data had `r attr(norm_data, "data_info")$num_edata` reads and `r attr(norm_data, "data_info")$num_samps` samples. There were `r attr(norm_data, "data_info")$num_miss_obs` instances of missing data (i.e., `r attr(norm_data, "data_info")$frac_missing * 100`% missing). 

There were `r length(unique(attr(norm_data,"group_DF")$Group))` groupings of these samples. The samples were spread among the groups as follows.

```{r, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, comment=""}
table(attr(norm_data,"group_DF")$Group)
```

```{asis, echo=ifelse(is.null(attr(norm_data,"data_info")$norm_method),FALSE,TRUE)}
## Normalization

The data was normalized using the `r attr(norm_data, "data_info")$norm_method` method.
```

```{asis, echo=ifelse(is.null(attr(norm_data, "filters")), FALSE, TRUE)}
## Filters
```

```{r, eval=ifelse(is.null(attr(norm_data, "filters")), FALSE, TRUE)}
unname(unlist(lapply(1:length(names(attr(norm_data, "filters"))), function(x) attr(norm_data, "filters")[[x]][1])))
```

Finally, data was distributed as follows.

```{r}
plot(norm_data)
```

```{r}
community <- any(c(!is.null(data$rich), !is.null(data$abun), !is.null(data$adiv), !is.null(data$even), !is.null(data$jaccard)))
```

```{asis, echo=ifelse(community, TRUE, FALSE)}
# Community Metrics

The following is information about community metrics that were run on the data.

```

```{asis, echo=ifelse(!is.null(data$rich), TRUE, FALSE)}
## Richness

Richness is the number of unique reads seen in a sample. This can be in the form of the observed number of reads or in the form of a richness estimator (Chao1 or ACE), which estimates the total number of species present in a community.

```

```{r, eval=ifelse(!is.null(data$rich), TRUE, FALSE)}
summary(data$rich)

plot(data$rich)
```

```{asis, echo=ifelse(!is.null(data$abun), TRUE, FALSE)}
## Abundance

Abundance is the total number of reads seen in a sample. This can vary drastically among samples for a variety of reasons, including ecological as well as technical variability. Normalization techniques attempt to lessen the differences between samples.

```

```{r, eval=ifelse(!is.null(data$abun), TRUE, FALSE)}
summary(data$abun)

plot(data$abun)
```

```{asis, echo=ifelse(!is.null(data$rich) & !is.null(data$abun), TRUE, FALSE)}
## Richness vs Abundance

A plot showing richness versus abundance can help determine how well a normalization strategy worked. Normalization should decrease the correlation between richness and abundance.

```

```{r, eval=ifelse(!is.null(data$rich) & !is.null(data$abun), TRUE, FALSE)}
plot(data$rich, data$abun)
```

```{asis, echo=ifelse(!is.null(data$adiv), TRUE, FALSE)}
## Alpha Diversity

Alpha diversity is a measure of how diverse a sample is.

```

```{r, eval=ifelse(!is.null(data$adiv), TRUE, FALSE)}
summary(data$adiv)

plot(data$adiv)
```

```{asis, echo=ifelse(!is.null(data$even), TRUE, FALSE)}
## Evenness

Evenness is a measure of the spread of reads in a sample. For instance, a sample with reads of equal abundance will have a high evenness value, whereas a sample where there is a large variance in the abundance of reads will have a lower evenness.

```

```{r, eval=ifelse(!is.null(data$even), TRUE, FALSE)}
summary(data$even)

plot(data$even)
```

```{asis, echo=ifelse(!is.null(data$jaccard), TRUE, FALSE)}
## Jaccard `r ifelse(attributes(data$jaccard)$similarity, "Similarity", "Dissimilarity")`

The Jaccard index is a binary beta diversity measure. Here, it is implemented to look at the `r ifelse(attributes(data$jaccard)$similarity, "similarity", "dissimilarity")` between samples in the same group.

```

```{r, eval=ifelse(!is.null(data$jaccard), TRUE, FALSE)}
summary(data$jaccard)

plot(data$jaccard)
```

```{asis, echo=ifelse(!is.null(data$effsp), TRUE, FALSE)}
## Effective Species

```

```{r, eval=ifelse(!is.null(data$effsp), TRUE, FALSE)}
summary(data$effsp)

plot(data$effsp)
```

```{asis, echo=ifelse(!is.null(data$diffabun), TRUE, FALSE)}
# Differential Abundance

Differential abundance analysis performs statistical pairwise comparisons to determine which reads are statistically different between the groups. This results in a p-value and log-fold-change for every read in the two groups.

```

```{r, eval=ifelse(!is.null(data$diffabun), TRUE, FALSE)}
diffexp_data <- data$diffabun

attr(diffexp_data,"Tests")$Text
```

```{r, eval=ifelse(!is.null(data$diffabun), TRUE, FALSE)}
summary(diffexp_data)
```

```{asis, echo=ifelse(!is.null(data$diffabun), TRUE, FALSE)}
The number and regulatory direction of significantly differentially abundant reads (using a p-value threshold of `r attr(results,"Threshold")` for each comparison and each test performed can be seen in the following plots. 

```

```{r, eval=ifelse(!is.null(data$diffabun), TRUE, FALSE)}
plot(diffexp_data, type="flag")
```

```{asis, echo=ifelse(!is.null(data$diffabun), TRUE, FALSE)}
The fold changes are plotted against the p-values in the following volcano plots.

```

```{r, eval=ifelse(!is.null(data$diffabun), TRUE, FALSE)}
plot(diffexp_data, type="volcano")
```

```{asis, echo=ifelse(!is.null(data$diffabun), TRUE, FALSE)}
The fold changes for significant features are plotted in the following heatmaps.

```

```{r, eval=ifelse(!is.null(data$diffabun), TRUE, FALSE)}
plot(diffexp_data, type="logfc")
```

```{asis, echo=ifelse(!is.null(data$diffabun), TRUE, FALSE)}
The significant features can be seen in the following plots.

```

```{r, eval=ifelse(!is.null(data$diffabun), TRUE, FALSE)}
plot_significant(diffexp_data, norm_data)
```

```{asis, echo=ifelse(!is.null(data$diffabun), TRUE, FALSE)}
The taxonomic distribution of the significantly differentially abundant features can be seen in the following plot(s).

```

```{r, eval=ifelse(!is.null(data$diffabun), TRUE, FALSE)}
plot_all_diffabun(diffexp_data, norm_data)
```

```{asis, echo=ifelse(!is.null(data$indsp), TRUE, FALSE)}
# Indicator Species

Indicator species are species which are characteristic of a group.

```

```{r, eval=ifelse(!is.null(data$indsp), TRUE, FALSE)}
summary(data$indsp)

plot(data$indsp, type="flag")
```

```{asis, echo=ifelse(!is.null(data$combined), TRUE, FALSE)}
# Combined

These are the features that were found significant in both the differential abundance tests and the indicator species analysis.

```

```{r, eval=ifelse(!is.null(data$combined), TRUE, FALSE)}
data$combined
```